---
title: "The Development of Syntagmatic Redundancy in ME"
author: "Raquel Montero"
date: "12.09.2023"
output: 
  html_document: 
      toc: true
      toc_float:  true
      number_sections: true
      theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
```
```{css, echo=FALSE}
h1,h4 {
  text-align: center;
}
p {
  text-align: justify; 
}

```
# Introduction
This file shows the code used to analyse the data in Section 3  of the paper "Sociolinguistic typology meets historical corpus linguistics"; submitted to the Transactions of the Philological Society. Information on how the data was searched, annotated etc. can be found in the Github repository. The following is a guide to the documents that the repository contains:


- Annotation Guidelines: AnnotationGuidelinesSyntRedME.pdf
- CorpusSearch: CorpusSearchSyntRedME_Adjectives.c//CorpusSearchSyntRedME_Quantifiers.c
- Geographical Information about the texts: GeographicsSyntRedME.pdf
- Final data: SyntRedME_Adjectives.cod.ooo//SyntRedME_Quantifiers.cod.ooo


# Data Cleaning
First, we need to load the packages that will be used: 
```{r}
library(gdata, warn.conflicts = FALSE)        # for drop-levels
library(plyr, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(ggrepel, warn.conflicts = FALSE)
library(splines, warn.conflicts = FALSE)
library(MASS, warn.conflicts = FALSE)
library(reshape2, warn.conflicts = FALSE)
library(maps, warn.conflicts = FALSE)
library(viridisLite, warn.conflicts = FALSE)
library(viridis, warn.conflicts = FALSE)       # colors of the map
library(DHARMa, warn.conflicts = FALSE)       # for residuals
library(colorspace)

```

Then, we import the annotated data, and prepare it for visualization (e.g. eliminating variables that are not needed, converting to numeric the scores, etc.):

```{r}
## ADJECTIVES DATA:


# Importing the data:

data_agreementadj <- read.delim("C:/Users/raque/OneDrive/Documents/2023PHD/PhD2/Papers/Philological_society_STARFISH/Repository_CaseStudy_SyntagmaticRedundancyME/Data/SyntRedME_Adjectives.cod.ooo.ooo", header=F, sep=":") #This path has to be changed 

# Naming the columns:
colnames(data_agreementadj) <- c("position", 
                                 "agreement", 
                                 "Text", 
                                 "Year", 
                                 "Dialect", 
                                 "numsyll", 
                                 "animate")

# Eliminate variables not used: 
ex.dataadj <- subset(data_agreementadj,
                     position != "z" &                                
                       position != "multiple"&  # Eliminating cases with multiple adjectives
                       position != "Conjoined" &  # Eliminating cases of conjoined adjectives 
                       agreement!= "z"  & 
                       agreement!= "french-agree" &  # Eliminating cases with French  morphology
                       agreement != "adverbial-ly" & # Eliminating adjectives formed with -ly
                       agreement != "e-ending" &     # Eliminate adjectives with root ends in -e
                       agreement != "uninflected" & # Eliminate adjectives uninflected in OE
                       Text != "z"  & 
                       Year != "z" &
                       Dialect != "z"  &
                       numsyll == "monosyllabic"  # Only monosyllabic adjectives
                     )


## QUANTIFIERS DATA:

# Importing the data:
data_agreementquant <- read.delim("C:/Users/raque/OneDrive/Documents/2023PHD/PhD2/Papers/Philological_society_STARFISH/Repository_CaseStudy_SyntagmaticRedundancyME/Data/SyntRedME_Quantifiers.cod.ooo.ooo", header=F, sep=":")

# Naming the columns:
colnames(data_agreementquant) <- c("quant",
                                   "agreement", 
                                   "Text", 
                                   "Year", 
                                   "Dialect", 
                                   "numsyll")

# Eliminate variables not used: 
ex.dataquant <- subset(data_agreementquant, 
                         quant != "z" & 
                         agreement != "z" & 
                         Text != "z" & 
                         Year != "z"  & 
                         Dialect != "z"  & 
                         numsyll == "monosyllabic" 
                       )


# droplevels: to drop unused factor levels.
ex.dataadj <- droplevels(ex.dataadj)
ex.dataquant <- droplevels(ex.dataquant)


# years stored as numbers: 
ex.dataadj$Year <- as.numeric(as.character(ex.dataadj$Year))
ex.dataquant$Year <- as.numeric(as.character(ex.dataquant$Year))

# create a new column changing the agreement type into a binary variable: (agreement = 1, no agreement = 0)
ex.dataadj$agreement_new <- ifelse(ex.dataadj$agreement == "agreement", 1, 0)
ex.dataadj$agreement_new <- as.numeric(as.character(ex.dataadj$agreement_new)) # change into numberic value
ex.dataquant$agreement_new <- ifelse(ex.dataquant$agreement == "agreement", 1, 0)
ex.dataquant$agreement_new <- as.numeric(as.character(ex.dataquant$agreement_new)) #change into numeric value


```
 We then have a look at how much data there is for our data analysis:
```{r}

table <- table(ex.dataadj$Dialect)
table

table1 <- table(ex.dataquant$Dialect)
table1

```
Calculate the mean agreement per text

```{r}
plot.data2 <- ddply(ex.dataadj, .(Text,Dialect,Year),summarize, whet = mean(as.numeric(as.character(agreement_new)), na.rm = T), n = sum(!is.na(as.numeric(as.character(agreement_new)))))
plot.data2


plot.data4 <- ddply(ex.dataquant, .(Text,Dialect,Year),summarize, whet = mean(as.numeric(as.character(agreement_new)), na.rm = T), n = sum(!is.na(as.numeric(as.character(agreement_new)))))
head(plot.data4)
```
Merging the two data sets: 
```{r}
##First, merge the two data frames:
total <- merge(plot.data2,
               plot.data4,
               by=c("Text","Dialect","Year"))

colnames(total) <- c("Text", "Dialect","Year","adjectives", "nad", "quantifier", "nq")


##New column with the differences in the agreement proportion: 
total$diff <- total$adjectives - total$quantifier


```

#  Data Visualization

## Map

First we need to add a column with the coordinates for each text: 

```{r}
# Adding column based on new Coordinates: 

#Adding the Latitude: 

total4 <- total %>%
  mutate(Longitude = case_when(
    (Text== "medicinis") ~ "-0.56880000",
    (Text== "Norwich") ~ "-0.20312085",
    (Text== "Horses" ) ~ "-1.0318673",
    (Text=="Richard-rolle") ~ "-2.0014878",
    (Text=="Innocencium") ~ "-1.4743217",
    (Text=="Edthor") ~ "-2.2412109",
    (Text=="Mandeville") ~ "0.33441843",
    (Text=="Edmund-m4") ~ "0.0000000",
    (Text=="Gaytry") ~ "-2.2412109",
    (Text=="Capgrave-chrnicle") ~ "1.0000000", 
    (Text=="Capgrave-sermon") ~ "0.41325972",
    (Text=="Morte-a") ~ "-1.5536906",
    (Text=="Brut") ~ "-2.9340404",
    (Text=="Kentish-homilies") ~ "-1.8031714",
    (Text=="Royal") ~ "-0.80603102",
    (Text=="Polychronicon") ~ "-2.4498470",
    (Text=="Benrul") ~ "-1.0836749",
    (Text=="Chronicles") ~ "-0.24573507",
    (Text=="Gregory") ~ "-0.60829082",
    (Text=="Cloud") ~ "-1.3818890",
    (Text=="Mirk") ~ "-1.8076144",
    (Text=="Lambeth") ~ "-2.4116106",
    (Text=="Ancrene") ~ "-2.7064436",
    (Text=="Siege-j") ~ "-1.6660928",
    (Text=="Ayenbyte") ~ "1.0814800",
    (Text=="Edmund") ~ "-2.1569555",
    (Text=="Katherine-group") ~ "-2.7064436",
    (Text=="Boethious") ~ "-0.12772404",
    (Text=="Old-testament") ~ "-0.89464405",
    (Text=="Vices") ~ "0.12773182",
    (Text=="Equatorie") ~ "-0.67179476",
    (Text=="Hilton") ~ "-0.22077060",
    (Text=="Kentish-sermons") ~ "0.62063882",
    (Text=="Parson") ~ "-0.12772404",
    (Text=="Trinity") ~ "-1.6118882",
    (Text=="Melibee") ~ "-0.12772404",
    (Text=="Book-vices") ~ "0.11872816",
    (Text=="Astrolabe") ~ "-1.0318673",
    (Text=="wycliffite") ~ "-1.3150252",
    (Text=="Purvey") ~ "-0.12942373",
    (Text=="Reynard") ~ "0.0000000",
    (Text=="Aelred") ~ "-2.3237144",
    (Text=="Aelred-m4") ~ "0.0000000",
    (Text=="Reynes") ~ "1.5482236",
    (Text=="Psalter") ~ "0.0039963911",
    (Text=="New-testament") ~ "-0.27629328",
    (Text=="Kempe") ~ "1.0000000",
    
    ))


#Adding the Longitude: 

total5 <- total4 %>%
  mutate(Latitude = case_when(
    (Text== "medicinis") ~ "53.065628",
    (Text== "Horses" ) ~ "51.453486",
    (Text== "Norwich") ~ "-53.182303",
    (Text=="Richard-rolle") ~ "53.955093",
    (Text=="Innocencium") ~ "51.068150",
    (Text=="Edthor") ~ "54.690768",
    (Text=="Mandeville") ~ "52.376964",
    (Text=="Edmund-m4") ~ "52.505077",
    (Text=="Gaytry") ~ "54.690768",
    (Text=="Capgrave-chrnicle") ~ "52.666667",
    (Text=="Capgrave-sermon") ~ "52.744048",
    (Text=="Morte-a") ~ "52.321306",
    (Text=="Brut") ~ "52.018714",
    (Text=="Kentish-homilies") ~ "50.892692",
    (Text=="Royal") ~ "51.485871",
    (Text=="Polychronicon") ~ "51.688892",
    (Text=="Benrul") ~ "53.962009",
    (Text=="Chronicles") ~ "52.566821",
    (Text=="Gregory") ~ "51.357801",
    (Text=="Cloud") ~ "52.560221",
    (Text=="Mirk") ~ "53.029131",
    (Text=="Lambeth") ~ "52.255453",
    (Text=="Ancrene") ~ "52.370931",
    (Text=="Siege-j") ~ "52.039927",
    (Text=="Ayenbyte") ~ "51.280552",
    (Text=="Edmund") ~ "52.210949",
    (Text=="Katherine-group") ~ "52.370931",
    (Text=="Boethious") ~ "51.507407",
    (Text=="Old-testament") ~ "51.783499",
    (Text=="Vices") ~ "51.742427",
    (Text=="Equatorie") ~ "52.796561",
    (Text=="Hilton") ~ "52.161753",
    (Text=="Kentish-sermons") ~ "51.246812",
    (Text=="Parson") ~ "51.507407",
    (Text=="Trinity") ~ "51.563236",
    (Text=="Melibee") ~ "51.507407",
    (Text=="Book-vices") ~ "51.859491",
    (Text=="Astrolabe") ~ "51.507407",
    (Text=="wycliffite") ~ "53.063272",
    (Text=="Purvey") ~ "52.178413",
    (Text=="Reynard") ~ "52.505077",
    (Text=="Aelred") ~ "52.255726",
    (Text=="Aelred-m4") ~ "52.505077",
    (Text=="Reynes") ~ "52.632148",
    (Text=="Psalter") ~ "51.564752",
    (Text=="New-testament") ~ "52.162668",
    (Text=="Kempe") ~ "52.666667",
    ))


## The coordinates are stored as characters but we need the numbers: 

total5$Latitude <- as.numeric(as.character(total5$Latitude))
total5$Longitude <- as.numeric(as.character(total5$Longitude))


```

Create the map: 
```{r}
##Creating the map: 

worldmap = map_data('world')
knitr::kable(head(worldmap, 20))

ggplot() + 
  geom_polygon(data = worldmap, 
                        aes(x = long, y = lat, group = group), fill = 'gray80', color = 'black') + 
  coord_fixed(ratio = 1.3, 
              xlim = c(-6.5,1.5), 
              ylim = c(50.3, 59)) +  
  theme(
  panel.background = element_rect(fill = NA),
  panel.ontop = TRUE)
```

Add the information to the map: 
```{r}
worldmap = map_data('world')
knitr::kable(head(worldmap, 20))

mapengland2<- ggplot() + 
  geom_polygon(data = worldmap, aes(x = long, y = lat, group = group), fill = 'grey60', color = 'black') + 
  coord_fixed(ratio = 1.3, xlim = c(-7.5,1.5), ylim = c(50.3, 59)) +
  geom_point(data=total5, 
             aes(x=Longitude, 
                 y=Latitude, 
                 color=diff), 
             alpha=I(.7), 
             size=3.5, 
             position=position_jitter(h=0.10, w=0.10)) +  
    scale_color_continuous_sequential(palette = "GnBu",limits=c(-1,1) ) +
  theme(legend.position = c(0.87, 0.75))

mapengland2

```

## Barplots

```{r}
#Barplots by dialect and Text: 

barplot1 <-ggplot(data=total5, aes(x=Text, y=diff)) +
  geom_bar(stat="identity") +
  facet_grid(. ~ Dialect,
             scales = "free", 
             space = "free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

barplot1


#Individual plots
total5EM <-subset(total5,Dialect=="EM")
barplotEM <-ggplot(data=total5EM, 
                   aes(x=Text, y=diff)) +
  geom_bar(stat="identity",width=0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("Diff in Plural Agreement in Q and Adj: East Middlands") +
  ylab("Difference in Agreement in Q and Adj")+
  ylim(-1,1)
barplotEM

##############
total5S <-subset(total5,Dialect=="Southern")
barplotS <-ggplot(data=total5S, aes(x=Text, y=diff)) +
  geom_bar(stat="identity",width=0.5) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Diff in Plural Agreement in Q and Adj: South") +ylab("Difference in Agreement in Q and Adj")+ylim(-1,1)
barplotS
#############
total5WM <-subset(total5,Dialect=="WM")
barplotWM <-ggplot(data=total5WM, aes(x=Text, y=diff)) +
  geom_bar(stat="identity",width=0.5) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Diff in Plural Agreement in Q and Adj: West Middlands") +ylab("Difference in Agreement in Q and Adj")+ylim(-1,1)
barplotWM

total5N <-subset(total5,Dialect=="Northern")
barplotN <-ggplot(data=total5N, aes(x=Text, y=diff)) +
  geom_bar(stat="identity",width=0.5) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ggtitle("Diff in Plural Agreement in Q and Adj: North") +ylab("Difference in Agreement in Q and Adj")+ylim(-1,1)
barplotN
```
```{r}
totalm <- total5 %>%
  mutate(Period = case_when(
    (Text== "medicinis") ~ "M4",
    (Text== "Horses" ) ~ "M3",
    (Text=="Richard-rolle") ~ "M2",
    (Text== "Norwich") ~ "M3",
    (Text=="Innocencium") ~ "M4",
    (Text=="Edthor") ~ "M3",
    (Text=="Mandeville") ~ "M3",
    (Text=="Edmund-m4") ~ "M4",
    (Text=="Gaytry") ~ "M3",
    (Text=="Capgrave-chrnicle") ~ "M4",
    (Text=="Capgrave-sermon") ~ "M4",
    (Text=="Morte-a") ~ "M4",
    (Text=="Brut") ~ "M3",
    (Text=="Kentish-homilies") ~ "M1",
    (Text=="Royal") ~ "M3",
    (Text=="Polychronicon") ~ "M3",
    (Text=="Benrul") ~ "M3",
    (Text=="Chronicles") ~ "M1",
    (Text=="Gregory") ~ "M4",
    (Text=="Cloud") ~ "M3",
    (Text=="Mirk") ~ "M3",
    (Text=="Lambeth") ~ "M1",
    (Text=="Ancrene") ~ "M1",
    (Text=="Siege-j") ~ "M4",
    (Text=="Ayenbyte") ~ "M2",
    (Text=="Edmund") ~ "M3",
    (Text=="Katherine-group") ~ "M1",
    (Text=="Boethious") ~ "M3",
    (Text=="Old-testament") ~ "M3",
    (Text=="Vices") ~ "M1",
    (Text=="Equatorie") ~ "M3",
    (Text=="Hilton") ~ "M3",
    (Text=="Kentish-sermons") ~ "M2",
    (Text=="Parson") ~ "M3",
    (Text=="Trinity") ~ "M1",
    (Text=="Melibee") ~ "M3",
    (Text=="Book-vices") ~ "M3",
    (Text=="Astrolabe") ~ "M3",
    (Text=="wycliffite") ~ "M3",
    (Text=="Purvey") ~ "M3",
    (Text=="Reynard") ~ "M4",
    (Text=="Aelred") ~ "M2",
    (Text=="Reynes") ~ "M4",
    (Text=="Psalter") ~ "M2",
    (Text=="New-testament") ~ "M3",
    (Text=="Kempe") ~ "M4",
    ))
```

# Statistical Analysis

## Linear Regression Model (lm)

```{r}
total$diff<- as.numeric(total$diff)

#Testing the effect of difference in agreement: linear model
ex.fittotal <- lm(diff~Dialect, data=total)
summary(ex.fittotal)

#using South as the reference level:
total$Dialect <- relevel(factor(total$Dialect), ref = 3)
ex.fittotal <- lm(diff~Dialect, data=total)
summary(ex.fittotal)

```
## Test of residuals (DHARMa)

```{r}
# DHARMa package: 

simulationOutput <- simulateResiduals(fittedModel = ex.fittotal, plot = F)
residuals(simulationOutput)
plotresiduals<- plot(simulationOutput)
plotresiduals


```






<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>





